Deterministic Scheduling of Periodic Messages in a Network

### 2017-07-13 article/article.tex
article

\begin{abstract}
We study the sending of periodic messages in a graph following predefined paths, a problem arising from communication between radio antennas and their signal processing units located in distant data-centers. Messages must be sent deterministically to avoid collisions on the network and to minimize the latency. We prove that problems arising from this context are $\NP$-complete even for very restricted graphs. We then deal with the case of a star graph and for different regimes of the parameters we propose algorithms that are experimentally shown to work extremely well compared to traditional statistical multiplexing.
\end{abstract}

@@

\todo{citation [7] pour les reseaux points a points ?}
 In this article, we study mechanisms to deal with the latency and periodicity constraints in the fronthaul network. 
 Statistical multiplexing even with a large bandwidth does not comply with the latency requirements. Therefore, the current solution~\cite{pizzinat2015things,tayq2017real} is to use dedicated circuits. Each end-point (RRH on one side, BBU on the other side) is connected through direct fiber or full optical switches. This architecture is very expensive and hardly scales in the case of a mobile network composed of about 10,000 base stations. Here we propose to use a deterministic approach to schedule our periodic messages. This approach has gained some traction recently: Deterministic Networking is under standardization in IEEE 802.1 TSN group~\cite{finn-detnet-architecture-08}, as well at IETF DetNet working group~\cite{ieee802}. Several patents on concepts and mechanisms for DetNet have been already published, see for example~\cite{howe2005time,leclerc2016transmission}. The aim is to operate a C-RAN on a low-cost shared switched network.

@@

      A {\bf $(P,\tau)$-periodic assignment} of a routed network $(G,\cal R)$ is a sequence  ${\cal M}=(m_1, \ldots ,m_n)$ of $n$ integers that we call {\bf offsets}, with $n$ the number of routes in $\cal R$. The number $m_{i}$ represents the time at which a message is emitted at the first vertex of the route $r_{i}$. We will always consider that the $m_{i}$ are between $0$ and $P-1$. Moreover, \emph{no pair of routes must have a collision} in a $(P,\tau)$-periodic assignment.

@@

      The following table summarize the main notations used in the paper.
      \begin{center}
   \begin{tabular}{|c|c|}
    \hline
     $(G,\cal R)$ & Routed network \\
     \hline
      $\omega(u,v)$ & Weight of the arc $(u,v) \in A$ \\
      \hline
      $\lambda(u_i,r)$ & Latency of the vertex $u_i$ in $r$\\
         \hline
         $\lambda(r)$ & Length of the route $r$\\
         \hline
         $P$ & Period\\
         \hline
         $\tau$ & Size of a message\\
         \hline
         $ [t(v,r)]$& Set of time slots used by route $r$ at vertex $v$ \\
         & in a period $P$\\
         \hline 
         ${\cal M}=(m_1, \ldots ,m_n)$& Assignment: an offset for each route)\\
              \hline 
         $w_r$& Waiting time of the route $r$\\
            \hline 
         $PT(r)$& Process time of the route $r$\\
           \hline 
         $MT({\cal M})$& Maximum process time of the assignment\\
    \hline

@@

  When the waiting times are zero, $MT({\cal M})$ is equal to twice the weight of the longest route, thus $T_{max}$ is not relevant anymore. For a route $r$, choosing the offset $m_r$ also sets the offset of the route $\rho(r)$ $m_{\rho(r)} + \lambda(r) \mod P$.
  Consider an assignment $\{m_0,\dots,m_{n-1}\}$ of a star topology and let $m_i'= m_{i} - t(c_s,r_i) \mod P$.
  Then $\{m_0',\dots,m_{n-1}'\}$ is an assignment of the same star topology where the weights of the arcs from $s_i$ to $c_s$ are $0$
  for all $i$. Therefore we assume from now on that the \emph{weight of the arcs from $s_i$ to $c_s$ are all equal to $0$}.

@@

      \begin{proposition} Let $(G, {\cal R})$ be a symmetric routed network, and let $n\tau + 2(\lambda(r_{n}) - \lambda(r_{0})) \leq P$, then there is $(P,\tau)$-periodic assignment of $(G, {\cal R})$ with all waiting times $0$ which is given by Shortest-Longest in time $O(n\log(n))$.\label{prop:SL}
      \end{proposition}
      \begin{proof}
       Since $m_{s_i} = i\tau$, $[t(c_s,r_{i})] = \{i\tau,\dots, (i+1)\tau -1\}$ and there are no collision on the forward period.

@@

        The offset of the route $r_i$ is set to $i\tau$, which ensures that there are no collision on the way forward (on the arc $(c_s,c_t)$) as soon as $P \geq n\tau$ which is the minimal possible period. We set the waiting time for the route $i$ to $w_i = 2(\lambda(r_{0}) - \lambda(r_{i}))$. We can compute the time at which the message of the route $r_i$ arrives at the vertex $c_t$ on its way back: $t(r_i,(c_t,c_s)) = w_i + i\tau + 2\lambda(r_{i})$
        by replacing $w_i$ by its value we obtain $t(r_i,(c_t,c_s)) =  i\tau + 2\lambda(r_{0})$
        As a conclusion there are no collision on the edge $(c_t,c_s)$ as soon as the 
        period is larger than $n\tau$ (there are no gaps between the messages).

@@

    \end{algorithmic}
    \end{algorithm}
    The function  min\_non\_assigned(eligible\_time) returns the lowest time in eligible\_time, for which the corresponding route is not assigned yet. The function update(free\_intervals) removes an interval of size $\tau$ beginning at t, which correspond to the message,  from free\_intervals.

@@

     We reduce our problem of setting the offsets of the backwards routes once the order of the forward routes is fixed to this
     scheduling problem. The backward routes are the jobs, the size of a message is the running time of a job,
     the deadline of a route is the deadline of the corresponding job and the smallest time at which it is eligible is the release time. Let us call {\bf Minimal Latency Scheduling (MLS)} this algorithm.
     Remark that we do not deal with the periodicity. When MLS finds a solution ${\cal M}$ its always satisfies $MT({\cal M}) < T_{max}$. On the other hand, it is a $(P,\tau)$ periodic assignment only if $m_n - m_1 \leq P -\tau$ where $m_n$ is the largest offset and $m_1$ the smallest one. The algorithm we use minimize $m_n - m_1 $, so it may work quite well in practice (as shown in Section~\ref{sec:resultsPALL}).

@@

     We have also computed the minimum margin needed by PMLS to find a solution but we have not represented it in Figure~\ref{fig:sto} since even the third quartile was always at $0$ for a period of $20000$.
     There is around $20\%$ of the instances for which the margin is larger than $0$ for $P=20000$ and $7\%$
     for $P = 21000$. For the instances with a margin larger than $0$, the margin is on average $400$.

>>>>

\begin{abstract}
We study the sending of periodic messages in a graph following predefined paths, a problem arising from communication between radio antennas and their signal processing units located in distant data-centers. Messages must be sent deterministically to avoid collisions on the network and to minimize the latency. We prove that problems arising from this context are $\NP$-complete even for very restricted graphs. We then deal with the case of a star graph and for different regimes of parameters we propose algorithms that are experimentally shown to work extremely well compared to traditional statistical multiplexing.
\end{abstract}

@@

\todo{citation [7] pour les reseaux points a points ?}
 In this article, we study mechanisms to deal with the latency and periodicity constraints in the fronthaul network. \todo{The fronthaul network is the network located between the BUUs and the RRHs.
 Statistical multiplexing even with a large bandwidth does not comply with the latency requirements. Therefore, the current solution~\cite{pizzinat2015things,tayq2017real} is to use dedicated circuits. Each end-point (RRH on one side, BBU on the other side) is connected through direct fiber or full optical switches. This architecture is very expensive and hardly scales in the case of a mobile network composed of about 10,000 base stations. Here we propose to use a deterministic approach to schedule our periodic messages. This approach has gained some traction recently: Deterministic Networking is under standardization in IEEE 802.1 TSN group~\cite{finn-detnet-architecture-08}, as well at IETF DetNet working group~\cite{ieee802}. Several patents on concepts and mechanisms for DetNet have been already published, see for example~\cite{howe2005time,leclerc2016transmission}. The aim is to operate a C-RAN on a low-cost shared switched network.

@@

      A {\bf $(P,\tau)$-periodic assignment} of a routed network $(G,\cal R)$ is a sequence  ${\cal M}=(m_0, \ldots ,m_{n-1})$ of $n$ integers that we call {\bf offsets}, with $n$ the number of routes in $\cal R$. The number $m_{i}$ represents the time at which a message is emitted at the first vertex of the route $r_{i}$. We will always consider that the $m_{i}$ are between $0$ and $P-1$. Moreover, \emph{no pair of routes must have a collision} in a $(P,\tau)$-periodic assignment.

@@

      The following table summarize the main notations used in the paper.
      \begin{center}
   \begin{tabular}{|c|c|}
    \hline
     $(G,\cal R)$ & Routed network \\
     \hline
      $\omega(u,v)$ & Weight of the arc $(u,v) \in A$ \\
      \hline
      $\lambda(u_i,r)$ & Latency of the vertex $u_i$ in $r$\\
         \hline
         $\lambda(r)$ & Length of the route $r$\\
         \hline
         $P$ & Period\\
         \hline
         $\tau$ & Size of a message\\
         \hline
         $ [t(v,r)]$& Set of time slots used by route $r$ at vertex $v$ \\
         & in a period $P$\\
         \hline 
         ${\cal M}=(m_0, \ldots ,m_{n-1})$& Assignment: an offset for each route)\\
              \hline 
         $w_r$& Waiting time of the route $r$\\
            \hline 
         $PT(r)$& Process time of the route $r$\\
           \hline 
         $MT({\cal M})$& Maximum process time of the assignment\\
    \hline

@@

  When the waiting times are zero, $MT({\cal M})$ is equal to twice the lenght of the longest route, thus $T_{max}$ is not relevant anymore. For a route $r$, choosing the offset $m_r$ also sets the offset of the route $\rho(r)$ to $m_{r} + \lambda(r) \mod P$.
  Consider an assignment $\{m_0,\dots,m_{n-1}\}$ of a star topology and let $m_i'= m_{i} - t(c_s,r_i) \mod P$.
  Then $\{m_0',\dots,m_{n-1}'\}$ is an assignment of the same star topology where the weights of the arcs from $s_i$ to $c_s$ are $0$
  for all $i$. Therefore we assume from now on that the \emph{weight of the arcs from $s_i$ to $c_s$ are all equal to $0$}.

@@

      \begin{proposition} Let $(G, {\cal R})$ be a symmetric routed network, and let $n\tau + 2(\lambda(r_{n-1}) - \lambda(r_{0})) \leq P$, then there is $(P,\tau)$-periodic assignment of $(G, {\cal R})$ with all waiting times $0$ which is given by Shortest-Longest in time $O(n\log(n))$.\label{prop:SL}
      \end{proposition}
      \begin{proof}
       Since $m_{s_i} = i\tau$, $[t(c_s,r_{i})] = \{i\tau,\dots, (i+1)\tau -1\}$ and there are no collision on the forward period.

@@

        The offset of the route $r_i$ is set to $i\tau$, which ensures that there are no collision on the way forward (on the arc $(c_s,c_t)$) as soon as $P \geq n\tau$ which is the minimal possible period. We set the waiting time for the route $i$ to $w_i = 2(\lambda(r_{0}) - \lambda(r_{i}))$. We can compute the time at which the message of the route $r_i$ arrives at the vertex $c_t$ on its way back: $t((c_t,c_s),r_i) = w_i + i\tau + 2\lambda(r_{i})$
        by replacing $w_i$ by its value we obtain $t((c_t,c_s),r_i) =  i\tau + 2\lambda(r_{0})$
        As a conclusion there are no collision on the edge $(c_t,c_s)$ as soon as the 
        period is larger than $n\tau$ (there are no gaps between the messages).

@@

    \end{algorithmic}
    \end{algorithm}
    The function  min\_non\_assigned(eligible\_time) returns the lowest time in eligible\_time, for which the corresponding route is not assigned yet. The function update(t,free\_intervals) removes an interval of size $\tau$ beginning at t, which correspond to the message,  from free\_intervals.

@@

     We reduce our problem of setting the offsets of the backwards routes once the order of the forward routes is fixed to this
     scheduling problem. The backward routes are the jobs, the size of a message is the running time of a job,
     the deadline of a route is the deadline of the corresponding job and the smallest time at which it is eligible is the release time. Let us call {\bf Minimal Latency Scheduling (MLS)} this algorithm.
     Remark that we do not deal with the periodicity. When MLS finds a solution ${\cal M}$ its always satisfies $MT({\cal M}) < T_{max}$. On the other hand, it is a $(P,\tau)$ periodic assignment only if $m_{n-1} - m_0 \leq P -\tau$ where $m_{n-1}$ is the largest offset and $m_0$ the smallest one. The algorithm we use minimize $m_{n-1} - m_0 $, so it may work quite well in practice (as shown in Section~\ref{sec:resultsPALL}).

@@

     We have also computed the minimum margin needed by PMLS to find a solution but we have not represented it in Figure~\ref{fig:sto} since even the third quartile was always at $0$ for a period of $20000$.
     There is around $20\%$ of the instances for which the margin is larger than $0$ for $P=20000$ and $2\%$
     for $P = 21000$. For the instances with a margin larger than $0$, the margin is on average $400$.
